---
title: "Covid91 vaccine study Final2023"
author: "Leonel Bautista"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))

male = filter(FinalData, sex =="M")
female = filter(FinalData, sex =="F")
lgbtq = filter(FinalData, LGBTQ =="gay")
drugUser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbiage when referencing p-values in conclusions.

## Males

### Graphical Results   

```{r}
barchartGC(~infected + treatment,data=male)
barchartGC(~infected + treatment,data=male, type="percent")
```

Displayed above is a Bar chart illustrating the graphical outcomes exclusively for males sourced from the data set. The comparison involves the incidence of contracting Covid based on vaccination status. The left side represents individuals who contracted Covid, while the right side depicts men who did not.      

It is evident that the percentage of individuals in the placebo group who contracted Covid is more than twice that of those who remained unaffected. The group of individuals who did not contract Covid exhibits a much closer relationship. Considering the sample size, it is improbable that these findings occurred solely by chance.     

## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions