---
title: "Covid91 vaccine study Final2023"
author: "Leonel Bautista"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))

male = filter(FinalData, sex =="M")
female = filter(FinalData, sex =="F")
lgbtq = filter(FinalData, LGBTQ =="gay")
drugUser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbiage when referencing p-values in conclusions.

## Males

### Graphical Results   

```{r}
barchartGC(~infected + treatment,data=male)
barchartGC(~infected + treatment,data=male, type="percent")
```

Displayed above is a Bar chart illustrating the graphical outcomes exclusively for males sourced from the data set. The comparison involves the incidence of contracting Covid based on vaccination status. The left side represents individuals who contracted Covid, while the right side depicts men who did not.      

It is evident that the percentage of individuals in the placebo group who contracted Covid is more than twice that of those who remained unaffected. The group of individuals who did not contract Covid exhibits a much closer relationship. Considering the sample size, it is improbable that these findings occurred solely by chance.     

### Numerical Results   

```{r}
tableMale <- xtabs(~infected + treatment, data=male)
rowPerc(tableMale)
colPerc(tableMale)
```

The numeric findings corroborate the insights gleaned from the graphical data. The placebo group experienced Covid infection around 70% of the time, whereas the vaccinated group had a 30% incidence rate. It is evident that the vaccine demonstrates overall efficacy in this testing context.   

### Inferential Results   

```{r}
chisq.test(tableMale)
chisqtestGC(tableMale)
maleFisher <- fisher.test(tableMale)
maleFisher
```

The Chi-Squared test reveals a significantly low P value of $1.098\times10^{-13}$ for the analyzed data, well below the conventional threshold of 0.05. This minute value strongly suggests a notable difference in the data. Additionally, the Chi-Squared statistic stands at 55, considerably surpassing the one degree of freedom derived from this dataset. Notably, the likelihood of a non-vaccinated individual contracting Covid is 2.2 times higher than that of a vaccinated individual.   

### Confidence Interval   

```{r}
oddsMale = maleFisher$estimate/(maleFisher$estimate+1)*100
print(oddsMale)
maleLow=(maleFisher$conf.int[1]/(maleFisher$conf.int[1]+1))*100
maleHigh=(maleFisher$conf.int[2]/(maleFisher$conf.int[2]+1))*100
cat("\n","Condifence interval for Males: ",maleLow," to ",maleHigh)
```

## Females

### Graphical Results

```{r}
barchartGC(~infected + treatment,data=female)
barchartGC(~infected + treatment,data=female, type="percent")
```

Displayed here is a bar chart that illustrates the graphical outcomes exclusively for females derived from the data. The chart juxtaposes the rates of contracting Covid based on vaccination status, with the left side representing women who contracted Covid and the right side representing those who did not.   

It is evident that the percentage of individuals in the placebo group who contracted Covid is approximately one-third higher than the corresponding percentage in the non-contracted group. Conversely, the percentages of those who did not contract Covid are closely aligned.   

### Numerical Results

```{r}
tableFemale <- xtabs(~infected + treatment, data=female)
rowPerc(tableFemale)
colPerc(tableFemale)
```

The numeric findings align with the graphical outcomes, revealing that the placebo group experienced Covid contraction roughly 60% of the time, whereas the vaccinated group had a 40% contraction rate. This indicates the overall efficacy of the vaccine in this testing scenario, albeit with a lower effectiveness compared to male participants.   

### Inferential Results

```{r}
chisq.test(tableFemale)
chisqtestGC(tableFemale)
femaleFisher <- fisher.test(tableFemale)
femaleFisher
```

The results of the Chi-Squared test indicate that the P value for the analyzed data is less than the commonly accepted threshold of 0.05, specifically $2.732\times10^{-4}$. This low value suggests a significant difference in the data. Additionally, the Chi-Squared statistic is 13, surpassing the one degree of freedom associated with this dataset. Notably, the likelihood of a non-vaccinated woman contracting Covid is 1.5 times higher than that of a woman who has received the vaccine.   

### Confidence Interval

```{r}
oddsFemale = femaleFisher$estimate/(femaleFisher$estimate +1)*100
print(oddsFemale)
femaleLow=(femaleFisher$conf.int[1]/(femaleFisher$conf.int[1]+1))*100
femaleHigh=(femaleFisher$conf.int[2]/(femaleFisher$conf.int[2]+1))*100
cat("\n","Condifence interval for Females: ",femaleLow," to ",femaleHigh)
```

We see that the confidence interval is >50% and includes the 60% threshold, it is less effective here than in the mens group but still a good vaccine.   

## LGBTQ

### Graphical Results

```{r}
barchartGC(~infected + treatment,data=lgbtq)
barchartGC(~infected + treatment,data=lgbtq, type="percent")
```

Displayed above is a Bar chart that contrasts the visual outcomes exclusively among drug users based on collected data. The comparison focuses on the incidence of contracting Covid in relation to vaccination status, with the left side representing individuals who contracted Covid and the right side representing those who did not.   

It is evident that a significantly higher number of individuals in the vaccinated group experienced Covid contraction, whereas those who did not contract Covid exhibit a much closer correlation. Considering the sample size, the likelihood of this occurrence being purely coincidental is low.   

### Numerical Results

```{r}
tableLGBTQ <- xtabs(~infected + treatment, data=lgbtq)
rowPerc(tableLGBTQ)
colPerc(tableLGBTQ)
```

The numeric findings corroborate the insights gleaned from the graphical data. In the placebo group, Covid was contracted around 15% of the time, while the vaccinated group experienced Covid contraction at an 85% rate. This indicates that, contrary to expectations, the vaccine demonstrates ineffectiveness in this testing scenario. In fact, it appears to be inversely effective, potentially elevating the risk of Covid contraction within this group.   

### Inferential Results

```{r}
chisq.test(tableLGBTQ)
chisqtestGC(tableLGBTQ)
lgbtqFisher <- fisher.test(tableLGBTQ)
lgbtqFisher
```

The results of the Chi-Squared test indicate a significantly low P value, specifically $6.668\times10^{-8}$, well below the conventional threshold of 0.05. This small P value suggests a notable difference in the tested data. Furthermore, the Chi-Squared statistic stands at 29, exceeding the degrees of freedom (one) in this dataset. In this group, the likelihood of a non-vaccinated individual contracting Covid is approximately 1 in 5.    

### Confidence Interval

```{r}
oddsLGBTQ = lgbtqFisher$estimate/(lgbtqFisher$estimate +1)*100
print(oddsLGBTQ)
lgbtqLow=(lgbtqFisher$conf.int[1]/(lgbtqFisher$conf.int[1]+1))*100
lgbtqHigh=(lgbtqFisher$conf.int[2]/(lgbtqFisher$conf.int[2]+1))*100
cat("\n","Condifence interval for LGBTQ: ",lgbtqLow," to ",lgbtqHigh)
```

We see that the confidence interval is 7-26% far lower than 50% so we can say that this is not a very good vaccine.    

## Druggies


# Overall Results and Conclusions