---
title: "Covid91 vaccine study Final2023"
author: "Leonel Bautista"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))

male = filter(FinalData, sex =="M")
female = filter(FinalData, sex =="F")
lgbtq = filter(FinalData, LGBTQ =="gay")
drugUser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbiage when referencing p-values in conclusions.

## Males

### Graphical Results   

```{r}
barchartGC(~infected + treatment,data=male)
barchartGC(~infected + treatment,data=male, type="percent")
```

Displayed above is a Bar chart illustrating the graphical outcomes exclusively for males sourced from the data set. The comparison involves the incidence of contracting Covid based on vaccination status. The left side represents individuals who contracted Covid, while the right side depicts men who did not.      

It is evident that the percentage of individuals in the placebo group who contracted Covid is more than twice that of those who remained unaffected. The group of individuals who did not contract Covid exhibits a much closer relationship. Considering the sample size, it is improbable that these findings occurred solely by chance.     

### Numerical Results   

```{r}
tableMale <- xtabs(~infected + treatment, data=male)
rowPerc(tableMale)
colPerc(tableMale)
```

The numeric findings corroborate the insights gleaned from the graphical data. The placebo group experienced Covid infection around 70% of the time, whereas the vaccinated group had a 30% incidence rate. It is evident that the vaccine demonstrates overall efficacy in this testing context.   

### Inferential Results   

```{r}
chisq.test(tableMale)
chisqtestGC(tableMale)
maleFisher <- fisher.test(tableMale)
maleFisher
```

The Chi-Squared test reveals a significantly low P value of $1.098\times10^{-13}$ for the analyzed data, well below the conventional threshold of 0.05. This minute value strongly suggests a notable difference in the data. Additionally, the Chi-Squared statistic stands at 55, considerably surpassing the one degree of freedom derived from this dataset. Notably, the likelihood of a non-vaccinated individual contracting Covid is 2.2 times higher than that of a vaccinated individual.   

### Confidence Interval   

```{r}
oddsMale = maleFisher$estimate/(maleFisher$estimate+1)*100
print(oddsMale)
maleLow=(maleFisher$conf.int[1]/(maleFisher$conf.int[1]+1))*100
maleHigh=(maleFisher$conf.int[2]/(maleFisher$conf.int[2]+1))*100
cat("\n","Condifence interval for Males: ",maleLow," to ",maleHigh)
```

## Females

### Graphical Results

```{r}
barchartGC(~infected + treatment,data=female)
barchartGC(~infected + treatment,data=female, type="percent")
```


## LGBTQ



## Druggies


# Overall Results and Conclusions